Supporting Information for ``Nonparametric Block Bootstrap Kolmogorov--Smirnov Goodness-of-Fit Test''

### 2024-07-11 manuscript/supplement.tex
question on theta0 hat

\newcommand{\mc}[1]{\textcolor{blue}{MC: (#1)}}

@@

Given any $x_1,\dots,x_m$ in the domain of $F$, by the CLT for weakly dependent 
time series, we have the joint convergence of 
$V_n:=(G_n(x_1),\dots,G_n(x_m),\sqrt{n}(\hat{\theta}_n-\theta_0))$
to some Gaussian limit, with the covariance matrix being 
$\Sigma_{m+1}=\lim_{n\rightarrow+\infty}\text{Cov}(V_n)$.
Consider the empirical process $ G_n(x) = \sqrt{n} (F_n(x) - F(x)) $, where 
$ F_n(x) $ is the empirical distribution function based on $ X_1, \ldots, 
X_n $. To show finite-dimensional convergence, we need to establish that for 
any fixed $ x_1, \ldots, x_m $, the vector $ (G_n(x_1), \ldots, G_n(x_m)) $ 
converges in distribution to a multivariate normal distribution.
The empirical distribution function can be written as
$
F_n(x) = \frac{1}{n} \sum_{t=1}^n \mathbf{1}\{X_i \leq x\}.
$
Therefore, the empirical process is:
\[
G_n(x) = \sqrt{n} (F_n(x) - F(x)) = \frac{1}{\sqrt{n}} \sum_{i=1}^n 
(\mathbf{1}\{X_i \leq x\} - F(x)).
\]

To establish the asymptotic distribution of $ G_n(x) $, we consider the vector 
\[
(G_n(x_1), \ldots, G_n(x_m)) = \frac{1}{\sqrt{n}} \sum_{i=1}^n (\mathbf{1}\{X_i \leq x_1\} - 
F(x_1), \ldots, \mathbf{1}\{X_i \leq x_m\} - F(x_m)).
\]
We calculate the covariance structure of the limiting distribution. The 
covariance between $ G_n(x) $ and $ G_n(x^{\prime}) $ is:
\[
\text{Cov}(G_n(x), G_n(x^{\prime})) = \frac{1}{n} \sum_{t=1}^n \sum_{s=1}^n 
\text{Cov}(\mathbf{1}\{X_t \leq x\}, \mathbf{1}\{X_t \leq x^{\prime}\}).
\]
Since $ \{X_i\} $ is strictly stationary, the covariance depends only on the 
lag $ h = |t-s| $:
\[
\text{Cov}(\mathbf{1}\{X_t \leq x\}, \mathbf{1}\{X_t \leq x^{\prime}\}) = \text{Cov}(\mathbf{1}\{X_h \leq x\}, 
\mathbf{1}\{X_0 \leq x^{\prime}\}).
\]
Thus, the covariance can be expressed as
\[
\text{Cov}(G_n(x), G_n(x')) = \frac{1}{n} \sum_{t=1}^n \sum_{h=-t+1}^{n-t} 
\text{Cov}(\mathbf{1}\{X_h \leq x\}, 
\mathbf{1}\{X_0 \leq x^{\prime}\}).
\]

@@

\paragraph{Convergence in distribution}
Therefore, we have
\begin{align*}
\begin{pmatrix}
G_n(x)  \\ 
\sqrt{n}(F(x;\hat{\theta}_n)-F(x;\theta_0))
\end{pmatrix}\Rightarrow
\begin{pmatrix}
G_{\infty}(x) \\
f(x;\theta_0)U
\end{pmatrix}
\end{align*}
jointly, where $G_{\infty}(x)$ is a Gaussian process with the covariance 
function defined as
\begin{align*}
\text{Cov}(G_{\infty}(x),G_{\infty}(x'))=& 
\lim_{n\rightarrow+\infty}\frac{1}{n}\sum^{n}_{i,j=1}
\text{Cov}(\mathbf{1}\{X_i\leq x\},\mathbf{1}\{X_j\leq x'\})
\\=& \sum^{+\infty}_{h=-\infty}\text{Cov}(\mathbf{1}\{X_{h}\leq x\},
\mathbf{1}\{X_0\leq x'\}),
\end{align*}
and $U\sim N(0,s^2)$ with 
\begin{align*}
s^2=\lim_{n\rightarrow+\infty}\frac{1}{n}\sum^{n}_{i,j=1}\text{Cov}(h(X_i,
\theta_0),h(X_j;\theta_0))
= \sum^{+\infty}_{h=-\infty}\text{Cov}(h(X_{h},\theta_0),h(X_0;\theta_0)),
\end{align*}
and the covariance between $U$ and $G_{\infty}(x)$ is given by
\begin{align*}
\text{Cov}(U,G_\infty(x))=&\lim_{n\rightarrow+\infty}
\frac{1}{n}\sum^{n}_{i,j=1}\text{Cov}(h(X_i,\theta_0),\mathbf{1}\{X_j\leq x\})
\\=&\sum^{+\infty}_{h=-\infty}
\text{Cov}(h(X_{h},\theta_0),\mathbf{1}\{X_0\leq x\}).
\end{align*}
By the continuous mapping theorem, we obtain
\begin{align*}
T_n \overset{d}{\rightarrow} \sup_x |G_{\infty}(x)+f(x;\theta_0)U|.    
\end{align*}

@@

\begin{proof}[Proof of Proposition~2]
    We want to show that
\begin{align}\label{eq-con-boot}
\begin{pmatrix}
G_n^{(b)}(x)  \\ 
\sqrt{n}(F(x;\hat{\theta}_n^{(b)})-F(x;\hat{\theta}_0))
\end{pmatrix}\Rightarrow^{*}
\begin{pmatrix}
G_{\infty}(x) \\
f(x;\theta_0)U
\end{pmatrix}
\end{align}
where $\Rightarrow^*$ means convergence in distribution almost surely and 
$\hat{\theta}_0:=\mathbb{E}^{(b)}[\hat{\theta}_n^{(b)}]$. 
In the bootstrap world, we have an expansion similar to (\ref{eq-expan}):
\begin{align*}
\hat{\theta}_n^{(b)}-\hat{\theta}_0 = \frac{1}{n}\sum^{n}_{i=1}h(X_i^{(b)};
\hat{\theta}_0) + \frac{1}{\sqrt{n}}R_n^{(b)},    
\end{align*}
where $R_n^{(b)} \overset{p^*}{\rightarrow} 0$, where 
$\overset{p^*}{\rightarrow}$ means convergence in probability almost surely. 
Note that 
\begin{align*}
\mathbb{E}^{(b)}\left[\frac{1}{n}\sum^{n}_{i=1}h(X_i^{(b)};
\hat{\theta}_0)\right]
=&\frac{1}{n}\sum^{k}_{j=1}\mathbb{E}^{(b)}\left[\sum^{S_j+l-1}_{i=S_j}h(X_{i};
\hat{\theta}_0)\right]
\\=&\frac{k}{n^2}[\sum^{n - l + 1}_{j=1}\sum^{j+l-1}_{i=j}h(X_{i};
\hat{\theta}_0) +
\sum^{n}_{j=n - l + 2}[\sum^{n}_{i=j}h(X_{i};\hat{\theta}_0) +
\sum^{j - n + l -1}_{i=1}h(X_{i};\hat{\theta}_0)]]
\\=&\frac{k}{n^2}\sum^{n}_{i=1}l h(X_{i};\hat{\theta}_0).
\end{align*}

Moreover, we have
\begin{align*}
\text{var}^{(b)}\left(\frac{1}{\sqrt{n}}\sum^{n}_{i=1}h(X_i^{(b)};
\theta_0^{(b)})\right)
=&\text{var}^{(b)}\left(\frac{1}{\sqrt{n}}\sum^{k}_{j=1}\sum^{S_j+l-1}_{i=S_j}
h(X_{i};\hat{\theta}_0)\right)
\\=&\frac{1}{l}\text{var}^{(b)}\left(\sum^{S_1+l-1}_{i=S_1}h(X_{i};
\hat{\theta}_0)\right)
\\=&\frac{1}{ln}\sum^{n}_{t=1}(W_t-\bar{W})^2
\end{align*}
with
\[
  W_t=
 \begin{cases}
\sum^{t+l-1}_{i=t}h(X_i;\hat{\theta}_0)  & \text{if } t \leq n - 1 + 1,\\
\sum^{n}_{i=t}h(X_{i};\hat{\theta}_0) +
\sum^{t - n + l -1}_{i=1}h(X_{i};\hat{\theta}_0)  & \text{if } t \geq n - 1 + 2
\end{cases}
\]
and $\bar{W}=\sum^{n}_{t=1}W_t/n$,
which is expected to converge to $s^2$ almost surely.

@@

For $\sqrt{n}(F(x;\hat{\theta}_n^{(b)}) - F(x;\hat{\theta}_0))$, using the 
first-order Taylor 
expansion around $\hat{\theta}_0$ similarly to what was done in 
\citet{kunsch1989jackknife}, we have:
$
F(x;\hat{\theta}_n^{(b)}) \approx F(x;\hat{\theta}_0) + f(x; 
\hat{\theta}_0) (\hat{\theta}_n^{(b)} - \hat{\theta}_0),
$ so
$
\sqrt{n}(F(x;\hat{\theta}_n^{(b)}) - F(x;\hat{\theta}_0)) \approx \sqrt{n} 
(F(x;\hat{\theta}_0) + f(x; 
\hat{\theta}_0) (\hat{\theta}_n^{(b)} - \hat{\theta}_0) - F(x;\hat{\theta}_0)) 
\approx
f(x; 
\hat{\theta}_0) \sqrt{n} (\hat{\theta}_n^{(b)} - \hat{\theta}_0).
$

Substituting the expansion of $\hat{\theta}_n^{(b)}$:
$
\sqrt{n} (\hat{\theta}_n^{(b)} - \hat{\theta}_0) = \frac{1}{\sqrt{n}} 
\sum_{i=1}^{n} 
h(X_i^{(b)}; \hat{\theta}_0) + R_n^{(b)}.
$

Since $R_n^{(b)} \overset{p^{(b)}}{\rightarrow} 0$, 
$
\sqrt{n}(F(x;\hat{\theta}_n^{(b)}) - F(x;\hat{\theta}_0)) \approx 
f(x; \hat{\theta}_0) \left( \frac{1}{\sqrt{n}} \sum_{i=1}^{n} h(X_i^{(b)}; 
\hat{\theta}_0) \right).
$

The term $\frac{1}{\sqrt{n}} \sum_{i=1}^{n} h(X_i^{(b)}; \hat{\theta}_0)$ 
converges in distribution to a normal random variable $U$ with variance 
$s^2$.

>>>>

\newcommand{\mc}[1]{\textcolor{blue}{MC: (#1)}}
\newcommand{\jy}[1]{\textcolor{purple}{JY: (#1)}}

@@

Given any $x_1,\dots,x_m$ in the domain of $F$, by the CLT for weakly dependent 
time series, we have the joint convergence of 
$V_n:=(G_n(x_1),\dots,G_n(x_m),\sqrt{n}(\hat{\theta}_n-\theta_0))$
to some Gaussian limit, with the covariance matrix being 
$\Sigma_{m+1}=\lim_{n\rightarrow+\infty}\text{Cov}(V_n)$.
Consider the empirical process $ G_n(x) = \sqrt{n} (F_n(x) - F(x)) $, where 
$ F_n(x) $ is the empirical distribution function based on $ X_1, \ldots, 
X_n $. To show finite-dimensional convergence, we need to establish that for 
any fixed $ x_1, \ldots, x_m $, the vector $ (G_n(x_1), \ldots, G_n(x_m)) $ 
converges in distribution to a multivariate normal distribution.
The empirical distribution function can be written as
$
F_n(x) = \frac{1}{n} \sum_{t=1}^n \mathbf{1}\{X_i \leq x\}.
$
Therefore, the empirical process is
\[
G_n(x) = \sqrt{n} (F_n(x) - F(x)) = \frac{1}{\sqrt{n}} \sum_{i=1}^n 
(\mathbf{1}\{X_i \leq x\} - F(x)).
\]

To establish the asymptotic distribution of $ G_n(x) $, we consider the vector 
\[
  (G_n(x_1), \ldots, G_n(x_m)) =
  \frac{1}{\sqrt{n}} \sum_{i=1}^n
  \left(\mathbf{1}\{X_i \leq x_1\} - 
    F(x_1), \ldots, \mathbf{1}\{X_i \leq x_m\} - F(x_m)\right).
\]
We calculate the covariance structure of the limiting distribution. The 
covariance between $ G_n(x) $ and $ G_n(x^{\prime}) $ is:
\[
\text{Cov}(G_n(x), G_n(x^{\prime})) = \frac{1}{n} \sum_{t=1}^n \sum_{s=1}^n 
\text{Cov}\left(\mathbf{1}\{X_t \leq x\}, \mathbf{1}\{X_t \leq x^{\prime}\}\right).
\]
Since $ \{X_i\} $ is strictly stationary, the covariance depends only on the 
lag $ h = |t-s| $:
\[
\text{Cov}(\mathbf{1}\{X_t \leq x\}, \mathbf{1}\{X_t \leq x^{\prime}\}) = \text{Cov}(\mathbf{1}\{X_h \leq x\}, 
\mathbf{1}\{X_0 \leq x^{\prime}\}).
\]
Thus, the covariance can be expressed as
\[
\text{Cov}(G_n(x), G_n(x')) = \frac{1}{n} \sum_{t=1}^n \sum_{h=-t+1}^{n-t} 
\text{Cov}(\mathbf{1}\{X_h \leq x\}, 
\mathbf{1}\{X_0 \leq x^{\prime}\}).
\]

@@

\paragraph{Convergence in distribution}
Therefore, we have
\begin{align*}
\begin{pmatrix}
G_n(x)  \\ 
\sqrt{n}(F(x;\hat{\theta}_n)-F(x;\theta_0))
\end{pmatrix}\Rightarrow
\begin{pmatrix}
G_{\infty}(x) \\
f(x;\theta_0)U
\end{pmatrix}
\end{align*}
jointly, where $G_{\infty}(x)$ is a Gaussian process with the covariance 
function defined as
\begin{align*}
\text{Cov}(G_{\infty}(x),G_{\infty}(x'))=& 
\lim_{n\rightarrow+\infty}\frac{1}{n}\sum^{n}_{i,j=1}
\text{Cov}(\mathbf{1}\{X_i\leq x\},\mathbf{1}\{X_j\leq x'\})
\\=& \sum^{+\infty}_{h=-\infty}\text{Cov}(\mathbf{1}\{X_{h}\leq x\},
\mathbf{1}\{X_0\leq x'\}),
\end{align*}
and $U\sim N(0, s^2)$ with 
\begin{align*}
s^2=\lim_{n\rightarrow+\infty}\frac{1}{n}\sum^{n}_{i,j=1}\text{Cov}(h(X_i,
\theta_0),h(X_j;\theta_0))
= \sum^{+\infty}_{h=-\infty}\text{Cov}(h(X_{h},\theta_0),h(X_0;\theta_0)),
\end{align*}
and the covariance between $U$ and $G_{\infty}(x)$ is given by
\begin{align*}
\text{Cov}(U,G_\infty(x))=&\lim_{n\rightarrow+\infty}
\frac{1}{n}\sum^{n}_{i,j=1}\text{Cov}(h(X_i,\theta_0),\mathbf{1}\{X_j\leq x\})
\\=&\sum^{+\infty}_{h=-\infty}
\text{Cov}(h(X_{h},\theta_0),\mathbf{1}\{X_0\leq x\}).
\end{align*}
By the continuous mapping theorem, we obtain
\begin{align*}
T_n \overset{d}{\rightarrow} \sup_x |G_{\infty}(x)+f(x;\theta_0)U|.    
\end{align*}

@@

\begin{proof}[Proof of Proposition~2]
  We want to show that
  \jy{What is $\hat\theta_0$? Should it be $\hat\theta_n$?}
\begin{align}\label{eq-con-boot}
\begin{pmatrix}
G_n^{(b)}(x)  \\ 
\sqrt{n}(F(x;\hat{\theta}_n^{(b)})-F(x;\hat{\theta}_0))
\end{pmatrix}\Rightarrow^{*}
\begin{pmatrix}
G_{\infty}(x) \\
f(x;\theta_0)U
\end{pmatrix}
\end{align}
where $\Rightarrow^*$ means convergence in distribution almost surely and 
$\hat{\theta}_0:=\mathbb{E}^{(b)}[\hat{\theta}_n^{(b)}]$. 
In the bootstrap world, we have an expansion similar to (\ref{eq-expan}):
\begin{align*}
\hat{\theta}_n^{(b)}-\hat{\theta}_0 = \frac{1}{n}\sum^{n}_{i=1}h(X_i^{(b)};
\hat{\theta}_0) + \frac{1}{\sqrt{n}}R_n^{(b)},    
\end{align*}
where $R_n^{(b)} \overset{p^*}{\rightarrow} 0$, where 
$\overset{p^*}{\rightarrow}$ means convergence in probability almost surely. 
Note that 
\begin{align*}
\mathbb{E}^{(b)}\left[\frac{1}{n}\sum^{n}_{i=1}h(X_i^{(b)};
\hat{\theta}_0)\right]
=&\frac{1}{n}\sum^{k}_{j=1}\mathbb{E}^{(b)}\left[\sum^{S_j+l-1}_{i=S_j}h(X_{i};
\hat{\theta}_0)\right]
\\=&\frac{k}{n^2}[\sum^{n - l + 1}_{j=1}\sum^{j+l-1}_{i=j}h(X_{i};
\hat{\theta}_0) +
\sum^{n}_{j=n - l + 2}[\sum^{n}_{i=j}h(X_{i};\hat{\theta}_0) +
\sum^{j - n + l -1}_{i=1}h(X_{i};\hat{\theta}_0)]]
\\=&\frac{k}{n^2}\sum^{n}_{i=1}l h(X_{i};\hat{\theta}_0).
\end{align*}

Moreover, we have
\begin{align*}
\text{var}^{(b)}\left(\frac{1}{\sqrt{n}}\sum^{n}_{i=1}h(X_i^{(b)};
\theta_0^{(b)})\right)
=&\text{var}^{(b)}\left(\frac{1}{\sqrt{n}}\sum^{k}_{j=1}\sum^{S_j+l-1}_{i=S_j}
h(X_{i};\hat{\theta}_0)\right)
\\=&\frac{1}{l}\text{var}^{(b)}\left(\sum^{S_1+l-1}_{i=S_1}h(X_{i};
\hat{\theta}_0)\right)
\\=&\frac{1}{ln}\sum^{n}_{t=1}(W_t-\bar{W})^2
\end{align*}
with
\[
  W_t=
 \begin{cases}
\sum^{t+l-1}_{i=t}h(X_i;\hat{\theta}_0)  , & \text{if } t \leq n - 1 + 1,\\
\sum^{n}_{i=t}h(X_{i};\hat{\theta}_0) +
\sum^{t - n + l -1}_{i=1}h(X_{i};\hat{\theta}_0) , & \text{if } t \geq n - 1 + 2,
\end{cases}
\]
and $\bar{W}=\sum^{n}_{t=1}W_t/n$,
which is expected to converge to $s^2$ almost surely.

@@

For $\sqrt{n}(F(x;\hat{\theta}_n^{(b)}) - F(x;\hat{\theta}_0))$, using the 
first-order Taylor 
expansion around $\hat{\theta}_0$ similarly to what was done in 
\citet{kunsch1989jackknife}, we have:
$
F(x;\hat{\theta}_n^{(b)}) \approx F(x;\hat{\theta}_0) + f(x; 
\hat{\theta}_0) (\hat{\theta}_n^{(b)} - \hat{\theta}_0),
$ so
\begin{align*}
  \sqrt{n}(F(x;\hat{\theta}_n^{(b)}) - F(x;\hat{\theta}_0))
  &\approx \sqrt{n} (F(x;\hat{\theta}_0) +
  f(x; \hat{\theta}_0) (\hat{\theta}_n^{(b)} - \hat{\theta}_0) - F(x;\hat{\theta}_0)) \\
  &\approx  f(x; \hat{\theta}_0) \sqrt{n} (\hat{\theta}_n^{(b)} - \hat{\theta}_0).
\end{align*}

Substituting the expansion of $\hat{\theta}_n^{(b)}$:
\[
\sqrt{n} (\hat{\theta}_n^{(b)} - \hat{\theta}_0) = \frac{1}{\sqrt{n}} 
\sum_{i=1}^{n} 
h(X_i^{(b)}; \hat{\theta}_0) + R_n^{(b)}.
\]

Since $R_n^{(b)} \overset{p^{(b)}}{\rightarrow} 0$, 
\[
\sqrt{n}(F(x;\hat{\theta}_n^{(b)}) - F(x;\hat{\theta}_0)) \approx 
f(x; \hat{\theta}_0) \left( \frac{1}{\sqrt{n}} \sum_{i=1}^{n} h(X_i^{(b)}; 
\hat{\theta}_0) \right).
\]
The term $\frac{1}{\sqrt{n}} \sum_{i=1}^{n} h(X_i^{(b)}; \hat{\theta}_0)$ 
converges in distribution to a normal random variable $U$ with variance 
$s^2$.
